\myTitle

### 2014-08-15 tex/main-body-text.tex
Minor edits and tweaks.

\subsection*{{\sc flock}}
Here we translate the procedure given by \citeauthor{Duc&Tur2009} for the 
{\sc flock} algorithm into a specification in terms of the variables
defined in the previous section.  We provide the description for a case in which
it is assumed there are $K$ clusters.

@@

\section*{Conclusions}
In the past decade, the use of genetic markers to identify population structure and to attempt inference of
the number of genetic clusters ($K$) in a collection has dramatically increased in the fields of ecology, evolution, 
epidemiology, and conservation
genetics. The most commonly used program for doing this is {\sc structure} \citep{Pritchardetal2000,Falushetal2003}, 
a full-featured program with many options that is widely-used, well-tested, and universally regarded as a standard tool amongst molecular ecologists. 
We've shown here that {\sc flock} is essentially equivalent to the program {\sc structure} using the no-admixture model with non-correlated allele frequencies. In comparing results from {\sc flock} and {\sc structure} 
on a real data set from {\em O. mykiss} we find that the two programs give similar results, only, at values of
$K>4$, {\sc flock}'s results appear to be more variable.

Though we observed similar results
between the two programs at low values of $K$, {\sc flock}'s `best-run' solutions became somewhat inconsistent at higher values of $K$.
This observation was consistent with \citet{Duc&Tur2012}'s comparison of the 
two programs using simulated data (Fig. 4 in \citet{Duc&Tur2012})
for $K=4 \& 8, L=10$. 
This is likely a consequence of the rapidly increasing size of the search space over possible 
partitions as $K$ and $n$, the number of individuals, increase. 
The likelihood surface defined on partitions of this space will often
be multimodal---possessing numerous local peaks and troughs.
The peak that  {\sc flock}'s deterministic algorithm finds is entirely determined by the initial allocation of 
individuals, as it is a deterministic algorithm.
If the number of initial allocations does not increase
as $K$ is increased, the starting allocations become diffusely scattered in the space of all
possible partitions making it less likely that many (if any at all) will converge to the same local mode. 
This coupled with the fact that there is 
ambiguity in what defines a `best run' when all plateaus are of length zero complicates the interpretation
of results from {\sc flock} if no non-zero plateaus are encountered. The problem of getting 
caught in local modes is not unique to {\sc flock}---{\sc structure} also can converge to 
different solutions in complex spaces; however our results suggest this is a bigger problem for {\sc flock}.
By increasing the number of loci used in their simulations \citet{Duc&Tur2012} showed, surprisingly, that 
{\sc flock} did converge on $K=8$, but held no advantage in inference of $K$ over {\sc structure}. We suspect that this 
resulted from the confined search space defined
by the small number of individuals (n=240) used in the analysis. 
%They simulated data using 500 individuals per population but only used 30 of them in the analysis.
% if you have more loci than individuals per population with high Fst between groups it does great!

@@

While there has been lengthy 
discourse over the choice of estimators for $K$ in {\sc structure} 
\citep{Pritchardetal2000,Evannoetal2005,Wap&Gag2006,Gaoetal2011}, less has been written on how the number of 
individuals
and the number of starting allocations can affect not only the convergence of different runs of {\sc flock}, but also its 
inference of $K$. \citeauthor{Duc&Tur2012} developed estimation rules for $K$ based on simulated data 
sets with a modest number of individuals and different migration regimes and rates, numbers of loci, and $K$. 
By relying on the number of identical partitions (mean LLOD values) obtained by {\sc flock} over multiple
runs as a means of supporting 
a particular $K$,  similar---but not identical---partitions
will provide no support to a given $K$. Given {\sc flock}'s tendency to converge to 
different solutions with large $K$ and $n$, it seems that 
this method for estimating $K$ is unlikely to be reliable or workable in large, complex
problems.  One possible patch for this problem would be running
multiple small batches and using the programs {\sc clumpp} \citep{Jak&Ros2007} and
{\sc distruct} \citep{Rosenberg2004} to visually inspect the `best run' results, and then 
determine plateaus amongst {\sc flock}'s replicate solutions not by identical LLOD, but by visual inspection of 
partitions that
are very similar but are perhaps not identical; however such an approach would be labor-intensive
and would rely on arbitrary, difficult-to-reproduce decisions.  It probably better to use {\sc structure} in such cases, keeping in mind, of course, that
estimating $K$ is a hard problem and should only be undertaken with caution, regardless of the method 
used. This is especially true in real populations that do not conform to the assumptions of genetic clustering models.

We made our comparison between {\sc flock} and the {\em no-admixture} model in 
{\sc structure} which is not the default model used in {\sc structure}. Rather, by default, {\sc structure} uses 
its {\em with-admixture} model.  Such a model explicitly tries to account for the admixed origin
of individuals in the sample and thus may be a more appropriate model than the no-admixture model,
or {\sc flock}, when the sample contains individuals that are admixed between populations.  A further advantage
of the with-admixture model is that it allows inferences for the population of origin of individual alleles at 
different loci within individuals, and not just of the individuals themselves.   We also note that {\sc flock} 
corresponds to {\sc structure}'s model with uncorrelated allele frequency priors. \citet{Falushetal2003} found that
using their ``$F$-model,'' which assumes that allele frequencies between populations are correlated {\em a priori}, 
allowed {\sc structure} to dissect more subtle population structure than was possible using the prior with
uncorrelated allele frequencies.  Modifying {\sc flock} so as to reap similar benefits from using a prior
with correlated allele frequencies seems like it would be an interesting challenge, though difficult
because {\sc flock}'s formulation effectively integrates the allele frequencies out of the model.

>>>>

\subsection*{{\sc flock}}
Here we translate the procedure given by \citet{Duc&Tur2009} for the 
{\sc flock} algorithm into a specification in terms of the variables
defined in the previous section.  We provide the description for a case in which
it is assumed there are $K$ clusters.

@@

\section*{Conclusions}
In the past decade, the use of genetic markers to identify population structure and to attempt inference of
the number of genetic clusters ($K$) in a collection has dramatically increased in the fields of ecology, evolution, 
epidemiology, and conservation
genetics. The most commonly used program for doing this is {\sc structure} \citep{Pritchardetal2000,Falushetal2003}, 
a full-featured program with many options that is widely-used, well-tested, and universally regarded as a standard tool amongst molecular ecologists. 
We've shown here that the model incorporated by {\sc flock} is essentially equivalent to the no-admixture model with non-correlated allele frequencies within the program {\sc structure}. In comparing results from {\sc flock} and {\sc structure} 
on a real data set from {\em O. mykiss} we find that the two programs give similar results, only, at values of
$K>4$, {\sc flock}'s results appear to be more variable.

Though we observed similar results
between the two programs at low values of $K$, {\sc flock}'s `best-run' solutions became somewhat inconsistent at higher values of $K$.
This observation was consistent with \citet{Duc&Tur2012}'s comparison of the 
two programs using simulated data 
for $K=4$ \& 8, $L=10$ (Fig. 4 in \citet{Duc&Tur2012}). 
This is likely a consequence of the rapidly increasing size of the search space over possible 
partitions as $K$ and $n$, the number of individuals, increase. 
The likelihood surface defined on partitions of this space will often
be multimodal---possessing numerous local peaks and troughs.
The peak that  {\sc flock}'s deterministic algorithm finds is entirely determined by the initial allocation of 
individuals, as it is a deterministic algorithm.
If the number of initial allocations does not increase
as $K$ is increased, the starting allocations become diffusely scattered in the space of all
possible partitions making it less likely that many (if any at all) will converge to the same local mode. 
This coupled with the fact that there is 
ambiguity in what defines a `best run' when all plateaus are of length zero complicates the interpretation
of results from {\sc flock} if no non-zero plateaus are encountered. The problem of getting 
caught in local modes is not unique to {\sc flock}---{\sc structure} also can converge to 
different solutions in complex spaces; however our results suggest this is a bigger problem for {\sc flock}.
By increasing the number of loci used in their simulations, \citet{Duc&Tur2012} showed  that 
{\sc flock} was able to converge in a simulated data set with $K=8$, but held no advantage in inference of $K$ over {\sc structure}. That simulated data set, with $n=240$ was considerably smaller and less complex than the real data investigated here.
%They simulated data using 500 individuals per population but only used 30 of them in the analysis.
% if you have more loci than individuals per population with high Fst between groups it does great!

@@

While there has been lengthy 
discourse over the choice of estimators for $K$ in {\sc structure} 
\citep{Pritchardetal2000,Evannoetal2005,Wap&Gag2006,Gaoetal2011}, less has been written on how the number of 
individuals
and the number of starting allocations can affect not only the convergence of different runs of {\sc flock}, but also its 
inference of $K$. \citeauthor{Duc&Tur2012} developed estimation rules for $K$ based on simulated data 
sets with a modest number of individuals and different migration regimes and rates, numbers of loci, and $K$. 
By relying on the number of identical partitions (mean LLOD values) obtained by {\sc flock} over multiple
runs as a means of supporting 
a particular $K$,  similar---but not identical---partitions
will provide no support to a given $K$. Given {\sc flock}'s tendency to converge to 
different solutions with large $K$ and $n$, it seems that 
this method for estimating $K$ is unlikely to be reliable or workable in large, complex
problems.  One possible patch for this problem would be running
multiple small batches and using the programs {\sc clumpp} \citep{Jak&Ros2007} and
{\sc distruct} \citep{Rosenberg2004} to visually inspect the `best run' results, and then 
determine plateaus amongst {\sc flock}'s replicate solutions not by identical LLOD, but by visual inspection of 
partitions that
are very similar but are perhaps not identical; however such an approach would be labor-intensive
and would rely on arbitrary, difficult-to-reproduce decisions.  It is probably better to use {\sc structure} in such cases, keeping in mind, of course, that
estimating $K$ is a hard problem and should only be undertaken with caution, regardless of the method 
used. This is especially true in real populations which do not conform to the assumptions of genetic clustering models.

We made our comparison between {\sc flock} and the {\em no-admixture} model in 
{\sc structure} which is not the default model used in {\sc structure}. Rather, by default, {\sc structure} uses 
its {\em with-admixture} model.  Such a model explicitly tries to account for the admixed origin
of individuals in the sample and thus may be a more appropriate model than the no-admixture model,
or {\sc flock}, when the sample contains individuals that are admixed between populations.  A further advantage
of the with-admixture model is that it allows inference of the population of origin of individual alleles at 
different loci within individuals, and not just of the individuals themselves.   We also note that {\sc flock} 
corresponds to {\sc structure}'s model with uncorrelated allele frequency priors. \citet{Falushetal2003} found that
using their ``$F$-model,'' which assumes that allele frequencies between populations are correlated {\em a priori}, 
allowed {\sc structure} to dissect more subtle population structure than was possible using the prior with
uncorrelated allele frequencies.  Modifying {\sc flock} so as to reap similar benefits from using a prior
with correlated allele frequencies seems like it would be an interesting challenge, though difficult
because {\sc flock}'s formulation effectively integrates the allele frequencies out of the model.
